---
title: "Template"
author: "Nerissa"
date: "2025-01-20"
output: 
   html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    number_sections: false
---
  
<style type="text/css">
  
body{ /* Normal  */
  font-size: 12px;
  }
td {  /* Table  */
  font-size: 12px;
}
h1.title {
  font-size: 18px;
  color: DarkBlue;
}
h1 { /* Header 1 */
  font-size: 18px;
}
h2 { /* Header 2 */
  font-size: 18px;
}
h3 { /* Header 3 */
  font-size: 18px;
}
code.r{ /* Code block */
  font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
  font-size: 14px;
}
</style>

---



# Load the mtcars dataset
```{r}
data("mtcars")
head(mtcars)
```

To start a code chunk in R Markdown, use ```` ```{r} ```` and close it with ```` ``` ````. Chunk options could be added like this: ```` ```{r echo = FALSE} ````. There are two chunk options I think are important:  
 1. echo: Display the code, default is `TRUE`.   
 2. eval: Display the output, default is `TRUE`.
 
Here to create a line new in a list, double space at the end of the line above.
---

If comments are added at the top of the chunk, there should be empty lines above and below.

# Setting the RNG seed
```{r}
set.seed(123)
```

If comments are added at the bottom of the chunk, there should be empty lines above and below. There should be a space between `#` and the heading.

---

# Fit a linear regression model
```{r}
model <- lm(mpg ~ wt, data = mtcars)
summary(model)
```

---

# Plot the data and regression line
```{r}
plot(mtcars$wt, mtcars$mpg, 
     main = "Linear Regression: mpg vs. wt",
     xlab = "Weight (1000 lbs)", 
     ylab = "Miles per Gallon",
     pch = 16, col = "blue")
abline(model, col = "red", lwd = 2)
```

Go to Help > Cheat Sheets > R Markdown Cheat Sheet.